#https://ucrel.lancs.ac.uk/bnc2/bnc2guide.htm#m2prep

# Frequency
Frequencies are from the BNC. Frequency lists based on: Adam Kilgarriff
```{r}
frequency_df <- read_delim(here("01_data", "lemma.al.txt"), delim = " ", col_names = c("sort-order", "frequency", "word", "word-class"))
```

conj (conjunction)            34 
adv  (adverb)                427
v    (verb)                 1281
det  (determiner)             47
pron (pronoun)                46
interjection                  13
a    (adjective)            1124
n    (noun)                 3262
prep (preposition)            71
modal                         12
infinitive-marker              1

Function to map specific word classes to pos tags
```{r}
map_to_tags <- function(word_class) {
  case_mapping <- c(
    "conj"              = "CCONJ",
    "adv"               = "ADV",
    "v"                 = "VERB",
    "det"               = "DET",
    "pron"              = "PRON",
    "interjection"      = "INTJ",
    "a"                 = "ADJ",
    "n"                 = "NOUN",
    "prep"              = "ADP",
    "modal"             = "AUX",
    "infinitive-marker" = "SCONJ"
  )
  
    return(case_mapping[word_class])
}
```
  
Apply the mapping function and create a new column
```{r}
frequency_df <- frequency_df %>% mutate(pos_map = map_to_tags(`word-class`))
```


Lemma & pos and word-class + pos_map were group to get frequencies because some words occur multiple times but contain different word classes.

example: 
  3341 2139 above a
942 10719 above adv
786 12889 above prep

```{r}
frequency_df <- frequency_df %>% select("word", "frequency", "pos_map")

eye_table <- eye_table %>% mutate(lemma = tolower(lemma))
eye_table <- left_join(eye_table, frequency_df, by = c("lemma" = "word", "pos" = "pos_map")) %>%
  mutate(frequency = replace(frequency, is.na(frequency), 0))
```
*Note* Curiously, spacey tags determiners as DT with a pos of PRON, which leads to a null value in the frequency cell because in *frequency_df*, determiners are mapped to determiners. 
Investigate if there are other such cases. -> there are, the respective cases are extracted below. There are some cases, where pos tags from the frequency table diverge from the ones in the main table (*case 1*) and some instances, where there exist multiple entries with different pos tags (*case 2*). For the second case, the frequency will be extracted manually.

```{r get_all_unique_missing_frequencies, eval=FALSE, include=FALSE}
unidentified_lemmas <- unique(eye_table$lemma[which(eye_table$frequency == 0)])
not_matched <- subset(frequency_df, frequency_df$word %in% unidentified_lemmas)
#print(not_matched)
```

```{r get_existing_missing_frequencies, eval=FALSE, include=FALSE}
subset_eye_table <- subset(eye_table, frequency == 0)
# Get unique combinations of lemma and pos
unique_lemmas_pos <- unique(subset_eye_table[, c("lemma", "pos")])
# Keep only rows where "lemma" occurs in the "word" column of not_matched
filtered_unique_lemma_pos <- unique_lemmas_pos[unique_lemmas_pos$lemma %in% not_matched$word, ]
# Order
ordered_filtered_unique_lemma_pos <- filtered_unique_lemma_pos[order(filtered_unique_lemma_pos$lemma), ]
```
```{r}
unidentified_lemmas <- unique(eye_table$token[which(eye_table$Frequency == 0)])
not_matched <- subset(frequency_spoken, frequency_spoken$Word %in% unidentified_lemmas)
print(not_matched)
```

```{r get_case_1_frequencies, eval=FALSE, include=FALSE}
# Identify unique entries in the "word" column
unique_words <- not_matched$word[!duplicated(not_matched$word) & !duplicated(not_matched$word, fromLast = TRUE)]
# Create a new dataframe with rows where "word" is unique
unique_not_matched <- not_matched[not_matched$word %in% unique_words, ]
```

```{r combine_missing_freq_cases, eval=FALSE, include=FALSE}
# Perform anti-join based on the "word" and "lemma" columns
result <- anti_join(ordered_filtered_unique_lemma_pos, unique_not_matched, by = c("lemma" = "word"))
# Add a new column "frequency" with values 0 to the result dataframe
result <- mutate(result, frequency = 0)
merged_table <- full_join(unique_not_matched, result, by = c("word" = "lemma", "pos_map" = "pos", "frequency"))
```

```
# List of words to be kept
words_to_keep <- c("all", "both", "fast", "kind", "like", "most", "neither", "own", "that", "this", "though", "to", "when", "where", "while", "worth")

# Filter rows where word is in the words_to_keep list
filtered_not_matched <- not_matched %>%
  filter(word %in% words_to_keep)

# Print or use the filtered_not_matched dataframe as needed
print(filtered_not_matched)
```

Manually mapped:
  all	PRON 230737	
both	PRON 41162	
fast	VERB 5527		
kind	VERB 1152		
like	INTJ	110090		
most	ADJ	54966		
neither	CCONJ	3018		
own	ADJ	6236		
that	PRON	384313		
this	PRON 461945
though	SCONJ	34801		
to	PART	1039323		
when	SCONJ	113655		
where	SCONJ	44496		
while	SCONJ	50548		
worth	ADJ 9183

```{r manually_input_frequencies, eval=FALSE, include=FALSE}
# Define a data frame with the specified frequencies for certain words
custom_frequencies <- data.frame(
  word = c("all", "both", "fast", "kind", "like", "most", "neither", "own", "that", "this", "though", "to", "when", "where", "while", "worth"),
  frequency = c(230737, 41162, 5527, 1152, 110090, 54966, 3018, 6236, 384313, 461945, 34801, 1039323, 113655, 44496, 50548, 9183)
)

# Merge the custom_frequencies data frame with merged_table based on the "word" column
modified_merged_table <- merge(merged_table, custom_frequencies, by = "word", all.x = TRUE)
# Replace the original frequency column with the custom frequencies where available
modified_merged_table$frequency <- ifelse(!is.na(modified_merged_table$frequency.y), modified_merged_table$frequency.y, modified_merged_table$frequency.x)
# Remove the unnecessary frequency columns
modified_merged_table <- select(modified_merged_table, -c(frequency.x, frequency.y))
print(modified_merged_table)
```

```{r merge_manual_frequencies_to_main, eval=FALSE, include=FALSE}
# Merge modified_merged_table with eye_table based on "word" = "lemma"
eye_table <- left_join(eye_table, modified_merged_table %>% select(word, frequency), by = c("lemma" = "word"))

# Replace the original frequency column with the merged frequency values
eye_table$frequency <- ifelse(!is.na(eye_table$frequency.y), eye_table$frequency.y, eye_table$frequency.x)

# Remove the unnecessary frequency columns
eye_table <- select(eye_table, -c(frequency.x, frequency.y))
```

```{r remove_freq_tables}
rm("unique_lemmas_pos", "unique_not_matched", "subset_eye_table", "result", "ordered_filtered_unique_lemma_pos", "not_matched", "modified_merged_table", "merged_table", "filtered_unique_lemma_pos", "custom_frequencies", "frequency_df")
```