# Statistical Exploration

**Libraries**
```{r message=FALSE, warning=FALSE}
library(lme4)
library(ggplot2)
library(hrbrthemes)
library(reshape2)
library(dplyr)
library(tidyr)
library(readr)
library(here)
```

## Dataset
```{r load_data, message=FALSE}
RT_data <- read_csv(here("05_output", "reading_times.csv"), col_names=TRUE, id=NULL)
```

```{r participant_count}
aggregate(participant_id ~ lang, data = RT_data, FUN = function(x) length(unique(x)))
```
```{r total_participants}
length(unique(RT_data$participant_id))
```

## Declare variables
```{r select_textnr}
#article <- RT_data %>% filter(textnr == 1)
article <- RT_data
```

"As regards very long fixations, deletion (but not merger) is also common. Eye-tracking researchers generally remove fixations longer than 800 ms, as recommended by Conklin and Pellicer-Sánchez (2016) and Keating (2014), because these fixations are believed to signal a lapse in attention." (Godfroid 2020: 261)
```{r outlier, eval=FALSE, include=FALSE}
# Cap the values in the dur column to a maximum of 800
#article$dur <- pmin(article$dur, 800)

# Remove duration over 800
#article <- article[article$dur <= 800, ]
```

### Log-Transformation
```{r log_duration}
article <- article %>% mutate(log_dur = log(dur)) # log transform
hist(article$dur)
hist(article$log_dur)
```
**Note:** Values that are *NA* must be dealt with. There are multiple options for handling missing values. 
While they could be removed, this would lead to a loss of data. Another approach is to set it to a very low value, for example 1ms.
```{r dependent_var}
article$dur[is.na(article$dur)] <- 0 
article$log_dur[is.na(article$log_dur)] <- 0

duration = article$dur
log_duration = article$log_dur
```

There are also AAPz scores for stopwords. The following code snippet changes all AAPz scores for stopwords to 0.
```{r independent_var}
# Update the AAPz column where is_stopword is 1
#article$AAPz[article$is_stopword == 1] <- 0

valency = article$AAPz
```

```{r categorical_valency}
# Turn into categorical
positive = 0.1
negative = -0.1

article <- article %>%
  mutate(AAPz_cat = case_when(
    AAPz > positive ~ "Positive",
    AAPz < negative ~ "Negative",
    TRUE ~ "Neutral"))
```

```{r confounding_var}
# lexical-level 
word_length = article$word_length
word_frequency = article$frequency
#predictability = TBD
POS = article$pos

# participant-level
participant = article$participant_id
```

How do I get bigrams?
```{r surprisal, eval=FALSE, include=FALSE}
bnc_size = 97626093

article <- article %>% 
  mutate(
    word = frequency,
    previous_word = lag(frequency),
    conditional_prob = 0,
    bigram_surprisal = log(1 / (word / bnc_size)) + log(1 / conditional_prob)
  )
```

```{r adjust_column}
article <- article %>% mutate(L1 = ifelse(lang == "en", "L1", "L2"))
article$exposure_EN[article$lang == "en"] <- 100 # set exposure for EN participants
```

```{r organize_columns}
article <- article %>% select(textnr, lang, L1, participant_id, 
                              tokenindex, token,
                              skip, pos, is_stopword,
                              dur, log_dur,
                              AAPz, AAPz_cat, anew_score, vader_score,
                              punctuation, word_length, frequency, 
                              motiv, lextale, cft20, 
                              start_learning,
                              exposure_EN
                              )

# fluent, reading, reading_fluent, exposure_L1,
# country, family, `school/workplace`,
# spelling, TOWRE_word, TOWRE_nonword, vocab, vocab.t2.5
```

## Means and Medians 
TODO: This doesn't say much because we need to take the other variables into account.
```{r}
article %>%
    group_by(AAPz_cat, lang) %>%
    summarise(val = median(dur)) %>% 
  pivot_wider(names_from = AAPz_cat, values_from = val)
```
#### Mean and median across all readers for each word
```{r calc_all}
article %>%
  group_by(token, tokenindex, textnr) %>%
  summarise(meanDurAll = round(mean(dur)), medianDurAll = round(median(dur)), .groups = 'drop')

article <- article %>%
    group_by(token, tokenindex, textnr) %>%
    mutate(meanDurAll = mean(dur), medianDurAll = median(dur), .groups = 'drop')
```
#### Median across language group for each word
Median duration of each token for each language group.
```{r calc_language}
median_duration <- article %>%
  group_by(textnr, token, tokenindex, lang) %>%
  summarise(medianDur = median(dur), .groups = 'drop') %>%
  spread(lang, medianDur, sep = "_") %>%
  rename_with(~gsub("^lang_", "median_", .), starts_with("lang_")) %>%
  rename(textnr = textnr, token = token, tokenindex = tokenindex)
```

```{r join_results}
article <- left_join(article, median_duration, by = c("textnr", "token", "tokenindex"))
rm(median_duration)
```

## Descriptive Statistics
#### Duration across language group 
```{r mean_lang}
article %>% group_by(lang) %>% 
  summarize(mean_duration = round(mean(dur)), 
            median_duration = median(dur), 
            standard_deviation = round(sd(dur)))
```

```{r plot_1}
ggplot(article, aes(x = lang, y = dur, fill = lang)) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 1000)) +
  scale_fill_manual(values = terrain.colors(9)) +
  labs(title = "Boxplot of Duration by Language",
       x = "Language",
       y = "Duration")
```

```{r mean_participant}
df2 <- article %>% group_by(participant_id) %>% summarize(mean_duration = mean(dur), L1) 
df2 <- unique(df2)
```

```{r}
cat("Slowest reader: ", 
    df2$participant_id[df2$mean_duration == max(df2$mean_duration)], ",",
    round(max(df2$mean_duration)),"ms", "\n")

cat("Fastest reader: ", 
    df2$participant_id[df2$mean_duration == min(df2$mean_duration)], ",",
    round(min(df2$mean_duration)), "ms")
```

The following density plot shows the mean reading times for each participant split between L1 and L2.
```{r density}
group_mean <- df2 %>%
  group_by(L1) %>%
  summarize(grp.mean = mean(mean_duration))

ggplot(df2, aes(x = mean_duration, fill = L1)) +
  geom_density(alpha = 0.5) +
  geom_vline(data=group_mean, aes(xintercept=grp.mean, color=L1),
             linetype="dashed") +
  labs(title = "Mean Reading Times between L1 and L2",
       x = "Duration",
       y = "Density") +
  theme_minimal()
```

## Correlation
**Note:** This uses the AAPz values. In this case, it must be determined how to deal with stopwords (see above).
```{r correlation}
cat("Correlation", "\n","Word length:", cor(word_length,duration), "\n","Word frequency:", cor(word_frequency, duration), "\n","Valency:", cor(valency, duration))
```

```{r correlation_by_participant}
df3 <- article %>%
  group_by(participant_id) %>%
  mutate(cor_valency=cor(AAPz, dur), cor_length=cor(word_length, dur), cor_frequency=cor(frequency, dur))

print(df3)
```

```{r corr_density}
# mean L1
mean_values <- df3 %>%
  group_by(L1) %>%
  summarize(mean_cor_valency = mean(cor_valency))

ggplot(df3, aes(x = cor_valency, fill = L1)) +
  geom_density(alpha = 0.5) +
  geom_vline(data = mean_values, aes(xintercept = mean_cor_valency, color = L1), linetype = "dashed") +
  labs(title = "Correlation between Valency and Duration by L1 and L2",
       x = "Correlation",
       y = "Density") +
  theme_minimal()
```

```{r correlation_by_language}
article %>%
  group_by(lang) %>%
  summarise(cor_valency=cor(AAPz, dur), cor_length=cor(word_length, dur), cor_frequency=cor(frequency, dur))
```

```{r}
ggplot(df3, aes(x = lang, y = cor_valency, fill = lang)) +
  geom_boxplot() +
  scale_fill_manual(values = terrain.colors(9)) +
  labs(title = "Boxplot of Correlation between Valency and Duration by Language",
       x = "Language",
       y = "Correlation")
```

Some exploratory median correlations:
```{r median_correlation}
cor(article$median_he, article$word_length)
cor(article$median_en, article$word_length)

cor(article$median_he, article$AAPz)
cor(article$median_en, article$AAPz)

cor(article$AAPz, article$frequency)
cor(article$AAPz, article$word_length)

summary(lm(median_ru ~ word_length + word_frequency + punctuation + valency, data = article))
```


## Linear Regression
```{r median_between}
article %>%
  group_by(token, tokenindex, L1) %>%
  summarise(valency = first(AAPz), medianDurAll = median(dur)) %>%
  pivot_wider(names_from = L1, values_from = medianDurAll)
```

```{r add_median}
article <- article %>%
  group_by(token, tokenindex, L1) %>%
  mutate(medianDurAll = median(dur), .after = dur)
```

```{r}
(sum(article$medianDurAll==0)/length(article$medianDurAll))*100
```
**important:** Median contains a lot of 0! mean performs better! However, need to check for the larger numbers.


multifactorial linear regression analysis 
- step-down and step-up analysis (also interactions between variables)
```{r RQ1}
summary(lm(medianDurAll ~ word_length + word_frequency + punctuation + valency + exposure_EN + 
             cft20 + lextale + motiv + start_learning, data = article))
```

Performs better without cft20!
```{r RQ2}
summary(lm(medianDurAll ~ word_length + word_frequency + punctuation + valency + 
             exposure_EN + cft20 + lextale + motiv + start_learning + lang + valency:lang, data = article))
```

```{r RQ3}
summary(lm(medianDurAll ~ L1 + word_length + word_frequency + punctuation + valency + valency + valency:L1, data = article))
```

```{r feature_weights}
weights <- lm(medianDurAll ~ word_length + word_frequency + punctuation + valency + exposure_EN + 
          cft20 + lextale + motiv + start_learning, data = article)

drop1(weights, test = "F")
```


## Linear Mixed-Models
Comment: Would need ANOVA to see which model is more significant
```{r eval=FALSE, include=FALSE}
summary(lmer(log_duration ~ valency + (1 + valency | word_length)))
summary(lmer(log_duration ~ valency + (1 + valency | participant)))
summary(lmer(log_duration ~ valency + (1 + valency | participant) + (1 | word_length)))
```
