# Statistical Exploration

## Libraries
```{r message=FALSE, warning=FALSE}
library(lme4)
library(ggplot2)
library(reshape2)
library(dplyr)
library(tidyr)
library(readr)
library(here)
```

## Declare variables
```{r load_data, message=FALSE}
reading_time <- read_csv(here("05_output", "reading_times.csv"), col_names=TRUE, id=NULL)
```

```{r select_textnr}
article <- reading_time %>% filter(textnr == 1)
article <- reading_time
```


```{r variables}
# dependent variable
duration = article$dur
log_duration = log(duration)

# independent variable
valency = article$AAPz

# confounders
# lexical-level 
word_length = article$word_length
word_frequency = article$frequency
#predictability = TBD
POS = article$pos

# participant-level
participant = article$participant_id
```

### Means and Medians 
#### Across all readers
```{r calc_all}
article$dur[is.na(article$dur)] <- 0
article %>%
  group_by(token, tokenindex) %>%
  summarise(meanDurAll = mean(dur), medianDurAll = median(dur))
```

```{r calc_language}
result <- article %>%
    group_by(tokenindex, lang, token) %>%
    summarise(medianDur = median(dur)) %>%
    spread(lang, medianDur, sep = "_")
```

```{r join_results}
median_article <- left_join(article, result, by = c("token", "tokenindex"))
```



## Correlation
**Note:** This uses the AAPz values. In this case, the stopwords also contain valency scores due to the AAPz. 
TODO: What is the correlation when the valency score for stopwords is set to 0?

```{r correlation}
duration[is.na(duration)] <- 0

cor(word_length,duration)
cor(word_frequency, duration)
cor(valency, duration)
```

with article set to **reading_time**
```
group_by(participant) %>%
+     summarize(cor=cor(valency, duration))
Error in UseMethod("group_by") : 
  no applicable method for 'group_by' applied to an object of class "factor"
```

```{r correlation_by_participant}
article$dur[is.na(article$dur)] <- 0

article %>%
  group_by(participant_id) %>%
  summarise(cor_valency=cor(AAPz, dur), cor_length=cor(word_length, dur), cor_frequency=cor(frequency, dur))
```
```{r correlation_by_language}
 cor_by_language <- article %>%
  group_by(lang) %>%
  summarise(cor_valency=cor(AAPz, dur), cor_length=cor(word_length, dur), cor_frequency=cor(frequency, dur))

print(cor_by_language)
```


Some exploratory median correlations:
```{r median_correlation}
cor(median_article$lang_he, median_article$word_length)
cor(median_article$lang_en, median_article$word_length)

cor(median_article$lang_he, median_article$AAPz)
cor(median_article$lang_en, median_article$AAPz)

cor(median_article$lang_he, median_article$frequency)
cor(median_article$lang_en, median_article$frequency)

cor(median_article$AAPz, median_article$frequency)
cor(median_article$AAPz, median_article$frequency)

cor(median_article$AAPz, median_article$word_length)
cor(median_article$AAPz, median_article$word_length)

lm(lang_ru ~ word_length + word_frequency + punctuation, data = median_article)
```


### Log-Transformation
Note: Values that are *NA* must be dealt with. There are multiple options for handling missing values. 
While they could be removed, this would lead to a loss of data.  
Another approach is to set it to a very low value, for example 1ms.
```{r}
hist(duration)
hist(log_duration)
```

## Regression
COMMENT: multifactorial linear regression analysis (multiple factors - see which one is significant)
- step-down and step-up analysis (also interactions between variables)
- begin with step-down 
- draw up a maximum model based on what is expected from the literature, 1 code per RQ -> send per email, then weed out the insignificant variables
```{r}
summary(lm(duration ~ valency))
```


```{r log_transform}
# Replace Inf and -Inf values with 0
log_duration[is.infinite(log_duration)] <- 0

summary(lm(log_duration ~ valency))
```
R^2 is 2.8%

# Linear Mixed-Models
Comment: Would need ANOVA to see which model is more significant
```{r}
summary(lmer(log_duration ~ valency + (1 + valency | word_length)))
```

```{r}
summary(lmer(log_duration ~ valency + (1 + valency | participant)))
```

```{r}
summary(lmer(log_duration ~ valency + (1 + valency | participant) + (1 | word_length)))
```



